\pkg{casebase}: An Alternative Framework For Survival Analysis and Comparison of Event Rates

### 2020-08-25 analysis/paper/paper.Rmd
added reason for smoothish cox and recolored

Figure \ref{fig:compAbsrisk} shows the absolute risk curves for both case-base sampling and the Fine-Gray model. As we can see, the two approaches agree quite well for AML; however, there seems to be a difference of about 5% between the two curves for ALL. This difference does not appear to be significant: the curve from case-base sampling is contained within a 95% confidence band around the Fine-Gray absolute risk curve (figure not shown).

# Case study 3---SUPPORT Data

@@

```{r supportData, eval = eval_cs3}
scaleFUNy <- function(x) sprintf("%.2f", x)
scaleFUNx <- function(x) sprintf("%.0f", x)
data(support)
# Change time to years
support$d.time <- support$d.time/365.25

@@

<!--Before generating an absolute risk curve using regularized estimation, we will explore three options as a baseline comparison. The original study compared these scores over a span of 180 days, which may not be comparable over the full 5.56 years. Therefore, we will examine three Cox regression scenarios: sps only, aps only and all covariates excluding sps and aps. Next, we will compare each modeling using the Brier score [@graf1999ass]. The Brier score can demonstrate both discrimination and calibration in a single model [@graf1999ass] and provide a comparable metric between semi-parametric and parametric methods in the survival context. For all the following analyses in this case study, we will use 80% of the observations to train our models, while 20% will be used to generate absolute risk curves, all of which will be averaged to demonstrate each model's predictive ability. Over 5.56 years, \ref{fig:baselines} demonstrates that all the cox models have similar, if not identical average predictions. Figure \ref{fig:brier1} examines the brier scores as they change over time. As the full model has the lowest brier score on overall, it will be used in the following absolute risk comparison to regularized cox, regularized estimation and the unadjusted Kaplan-Meier curve.-->

@@

```{r baselines, echo=FALSE, eval=eval_cs3}
cox_ci <- ggplot(baseCoxComparisonData, aes(Time)) +
  geom_step(aes(y = APS, colour = "APS")) +
  geom_step(aes(y = Full, col = "Full")) +
  geom_step(aes(y = SPS, colour = "SPS")) +
  geom_step(data = abKMcompare, mapping = aes(x = Time, y = KM, color = "K-M")) +
  labs(y = "Probability of Death", x = "Follow-up time (Years)", color = "Models") +
  scale_color_manual(values = q4, breaks = c("APS", "Full", "SPS", "K-M")) + 
  paper_gg_theme+ scale_x_continuous(labels=scaleFUNx,breaks = round(seq(0, 5, by = 1))) +scale_y_continuous(labels=scaleFUNy,n.breaks=8)
```

@@

cox_brier <- ggplot(data = brierPlotter, mapping = aes(x = times, y = brierScore, col = model)) +
  geom_line() +
  xlab("Follow-up time (Years)") +
  ylab("Brier score") +
  labs(color = "Models") +
  # ggtitle("Brier scores over time") +
  scale_color_manual(values = q4, breaks = c("APS", "Full", "SPS", "K-M")) + 
  paper_gg_theme+ scale_x_continuous(labels=scaleFUNx,breaks = round(seq(0, 5, by = 1)))+scale_y_continuous(labels=scaleFUNy,n.breaks=8)

@@

For our penalized model, we opt for the natural log of time; recall that this corresponds to a Weibull distribution. For fitting the penalized hazard, we use \code{fitSmoothHazard.fit}, which is a matrix interface to the function \code{fitSmoothHazard}. In order to use this interface, we first create a matrix \code{y} containing the time and event variables, and a matrix \code{x} containing all remaining covariates we would like to include in the model space. We used lasso penalization by setting \code{alpha = 1}. Moreover, we use the argument \code{penalty.factor} to force the time variable into the final model.

```{r supportCB_fit, eval=eval_cs3, echo=TRUE}
x <- model.matrix(death ~ . - d.time - aps - sps, 
                  data = train)[, -c(1)] # Remove intercept
y <- data.matrix(subset(train, select = c(d.time, death)))

@@

Next, we can use the object \code{cbFull} to estimate the absolute risk curve on the test data. We first create a matrix \code{newx} containing the relevant variables. Next, using \code{times}, we specify the time points at which the cumulative incidence will be computed. Finally, using the parameter \code{s = "lambda.1se"}, we specify at which value of the tuning parameter we want to retrieve the coefficient estimates.

```{r supportCB_abs, eval=eval_cs3, echo=TRUE}
# Estimating the absolute risk curve using the newdata parameter
newx <- model.matrix(death ~ . - d.time - aps - sps, 
                  data = test)[, -c(1)]
times <- sort(unique(test$d.time))

@@

abCB <- data.frame(Time = abCbFull[, 1],
                   absRisk = rowMeans(abCbFull[, -c(1)]))
abCB$Model <- "Penalized estimation"

@@

Our first comparison plot examines the covariates that were selected between all three hazard functions. Note that Cox regression without penalization selected more covariates than the other two, but only the intersection of coefficients with values greater than 0 are displayed in \ref{fig:cs3lolliPlot}. A visible tendency is for Cox regression to have larger estimated coefficients. This is expected, as there are no restrictions on the size of coefficients. The smallest coefficient estimates are found in penalized case-base model, aside from cancer state (cano).

@@

estimatescb <- data.table::setDT(as.data.frame(coef(cbFull)[-c(1, 2), 1]), 
                     keep.rownames = TRUE)
# estimatescb$std.error<-0
estimatescb$Model <- "Pen. CB"
colnames(estimatescb) <- c("term", "estimate", "model")

@@

These values do not inform us of model performance. To do so, we take our hazard functions and use them to predict the absolute risk of everyone in our test set, over time. The probabilities over time for each individual are averaged, resulting in the absolute risk curves in Figure \ref{fig:cs3FinalBrier} (A). There is separation between the curves, but it is minimal. To better demonstrate the differences in performance between the models, we compare Brier scores over time. In Figure \ref{fig:cs3FinalBrier} (B), the Brier score is lower in our penalized case-base model compared to both Cox regression models. The differences, however, are minimal. All three outperform the Kaplan-Meier estimator.

```{r cs3Final, echo=FALSE, eval=eval_cs3}
reg_ci <- ggplot(ab, aes(Time)) +
  geom_step(data = baseCoxComparisonData, mapping = aes(x = Time,y = Full, colour = "Cox")) +
  geom_line(aes(y = absRisk, colour = Model)) +
  geom_step(data = abKMcompare, mapping = aes(x = Time, y = KM, color = "K-M")) +
  labs(y = "Probability of death", x = "Follow-up time (years)", color = "Models") +
  # ggtitle("Absolute risk curves") +
  scale_color_manual(values = q4, breaks = c("K-M","Cox","Penalized Cox","Penalized CB")) + 
  paper_gg_theme+ scale_x_continuous(labels=scaleFUNx,breaks = round(seq(0, 5, by = 1))) +scale_y_continuous(labels=scaleFUNy,n.breaks=9)
```

@@

```{r riskregressionBrier, echo = FALSE, eval = eval_cs3, fig.cap="\\label{fig:cs3FinalBrier} Comparison of Cox regression, penalized Cox regression, penalized case-base estimation, and Kaplan-Meier. (A)  Probabilities of death as it changes over follow-up time. (B) Brier scores as they change over time, where a lower score corresponds to better performace. Penalized case-base estimation produced the lowest results over follow-up time.", warning = FALSE}
# Compute Brier score for penalized models
brierFinalResults <- Score(list("Penalized Cox" = coxNet, "Penalized CB" = cbFull),
                           data = cbind(subset(test, select = c(d.time, death)),
                                        as.data.frame(newx)), 
                           formula = Hist(d.time, death != 0) ~ 1, summary = NULL, 
                           se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
#brierFinalResults$models<-as.character(brierFinalResults$models)
# Combine Brier scores between unpenalized and penalized models before plotting

data_brier <- bind_rows(
  brierFinalResults$Brier$score,
  filter(brierFull$Brier$score, model == "PredictionModel")
  ) %>% 
  mutate(model = as.character(model), # First return to string, so types match
         model = case_when(
    model == "Null model" ~ "K-M",
    model == "PredictionModel" ~ "Cox",
    TRUE ~ model
  ),
  model = factor(model, levels = c("Cox", "Penalized Cox", "Penalized Logistic", "K-M"),
                 labels = c("Cox", "Pen. Cox", "Pen. Logistic", "K-M")))

reg_brier <- ggplot(data = data_brier, mapping = aes(x = times, y = Brier, col = model)) +
  geom_line() +
  xlab("Follow-up time (years)") +
  ylab("Brier score") +
  # ggtitle("Brier score over time") +
  labs(color = "Models") +
  paper_gg_theme +  
  scale_color_manual(values = q4) + scale_x_continuous(labels=scaleFUNx,breaks = round(seq(0, 5, by = 1))) +scale_y_continuous(labels=scaleFUNy,n.breaks=8)

@@

>>>>

Figure \ref{fig:compAbsrisk} shows the absolute risk curves for both case-base sampling and the Fine-Gray model. As we can see, the two approaches agree quite well for AML; however, there seems to be a difference of about 5% between the two curves for ALL. This difference does not appear to be significant: the curve from case-base sampling is contained within a 95% confidence band around the Fine-Gray absolute risk curve (figure not shown).
# Case study 3---SUPPORT Data

@@

```{r supportData, eval = eval_cs3}
data(support)
# Change time to years
support$d.time <- support$d.time/365.25

@@

<!--Before generating an absolute risk curve using regularized logistic regression, we will explore three options as a baseline comparison. The original study compared these scores over a span of 180 days, which may not be comparable over the full 5.56 years. Therefore, we will examine three Cox regression scenarios: sps only, aps only and all covariates excluding sps and aps. Next, we will compare each modeling using the Brier score [@graf1999ass]. The Brier score can demonstrate both discrimination and calibration in a single model [@graf1999ass] and provide a comparable metric between semi-parametric and parametric methods in the survival context. For all the following analyses in this case study, we will use 80% of the observations to train our models, while 20% will be used to generate absolute risk curves, all of which will be averaged to demonstrate each model's predictive ability. Over 5.56 years, \ref{fig:baselines} demonstrates that all the cox models have similar, if not identical average predictions. Figure \ref{fig:brier1} examines the brier scores as they change over time. As the full model has the lowest brier score on overall, it will be used in the following absolute risk comparison to regularized cox, regularized logistic regression and the unadjusted Kaplan-Meier curve.-->

@@

```{r baselines, echo=FALSE, eval=eval_cs3}
cox_ci <- ggplot(baseCoxComparisonData, aes(Time)) +
  geom_step(aes(y = APS, colour = "APS")) +
  geom_step(aes(y = Full, col = "Full")) +
  geom_step(aes(y = SPS, colour = "SPS")) +
  geom_step(data = abKMcompare, mapping = aes(x = Time, y = KM, color = "K-M")) +
  labs(y = "Probability of Death", x = "Follow-up time (Years)", color = "Models") +
  scale_color_manual(values = q4, breaks = c("APS", "Full", "SPS", "K-M")) + 
  paper_gg_theme
```

@@

cox_brier <- ggplot(data = brierPlotter, mapping = aes(x = times, y = brierScore, col = model)) +
  geom_line() +
  xlab("Follow-up time (Years)") +
  ylab("Brier score") +
  labs(color = "Models") +
  # ggtitle("Brier scores over time") +
  scale_color_manual(values = q4, breaks = c("APS", "Full", "SPS", "K-M")) + 
  paper_gg_theme

@@

For our penalized logistic regression model, we opt for the natural log of time; recall that this corresponds to a Weibull distribution. For fitting the penalized hazard, we use \code{fitSmoothHazard.fit}, which is a matrix interface to the function \code{fitSmoothHazard}. In order to use this interface, we first create a matrix \code{y} containing the time and event variables, and a matrix \code{x} containing all remaining covariates we would like to include in the model space. We used lasso penalization by setting \code{alpha = 1}. Moreover, we use the argument \code{penalty.factor} to force the time variable into the final model.

```{r supportCB_fit, eval=eval_cs3, echo=TRUE, cache=FALSE}
x <- model.matrix(death ~ . - d.time - aps - sps, 
                  data = train)[, -c(1)] # Remove intercept
y <- data.matrix(subset(train, select = c(d.time, death)))

@@

Next, we can use the object \code{cbFull} to estimate the absolute risk curve on the test data. We first create a matrix \code{newx} containing the relevant variables. Using \code{times}, we specify the time points at which the cumulative incidence will be computed. Finally, using the parameter \code{s = "lambda.1se"}, we specify at which value of the tuning parameter we want to retrieve the coefficient estimates.

```{r supportCB_abs, eval=eval_cs3, echo=TRUE, cache=TRUE}
# Estimating the absolute risk curve using the newdata parameter
newx <- model.matrix(death ~ . - d.time - aps - sps, 
                  data = test)[, -c(1)]
times <- sort(unique(test$d.time))

@@

abCB <- data.frame(Time = abCbFull[, 1],
                   absRisk = rowMeans(abCbFull[, -c(1)]))
abCB$Model <- "Penalized logistic"

@@

Our first comparison plot examines the covariates that were selected between all three hazard functions. Note that Cox regression without penalization selected more covariates than the other two, but only the intersection of coefficients with values greater than 0 are displayed in \ref{fig:cs3lolliPlot}. A visible tendency is for Cox regression to have larger estimated coefficients. This is expected, as there are no restrictions on the size of coefficients. The smallest coefficient estimates are found in penalized logistic model, aside from cancer state (cano).

@@

estimatescb <- data.table::setDT(as.data.frame(coef(cbFull)[-c(1, 2), 1]), 
                     keep.rownames = TRUE)
# estimatescb$std.error<-0
estimatescb$Model <- "Pen. logistic"
colnames(estimatescb) <- c("term", "estimate", "model")

@@

```{r cs3Final, echo=FALSE, eval=eval_cs3}
reg_ci <- ggplot(ab, aes(Time)) +
  geom_step(data = baseCoxComparisonData, mapping = aes(x = Time,y = Full, colour = "Cox")) +
  geom_line(aes(y = absRisk, colour = Model)) +
  geom_step(data = abKMcompare, mapping = aes(x = Time, y = KM, color = "K-M")) +
  labs(y = "Probability of Death", x = "Follow-up time (Years)", color = "Models") +
  # ggtitle("Absolute risk curves") +
  scale_color_manual(values = q4, breaks = c("K-M","Cox","Penalized Cox","Penalized logistic")) + 
  paper_gg_theme
```

@@

```{r riskregressionBrier, echo = FALSE, eval = eval_cs3, fig.cap="\\label{fig:cs3FinalBrier} (A) Four curves comparing the probabilities of death as it changes over follow-up time between Cox regression, penalized Cox regression, penalized logistic regression, and Kaplan-Meier. All follow-up times seen in the new data were used to generate the curves. The Cox-based curves appear to be smooth, but this is due to the 1703 follow-up times seen while fitting the models and the 319 time points used to estimate the curves. (B) Four curves demonstrating the Brier scores as they change over time between Cox regression, penalized Cox regression, penalized logistic regression, and Kaplan-Meier. Lower is better. Penalized logistic regression produced the lowest Brier scores over follow-up time.", warning = FALSE}
# Compute Brier score for penalized models
brierFinalResults <- Score(list("Penalized Cox" = coxNet, "Penalized Logistic" = cbFull),
                           data = cbind(subset(test, select = c(d.time, death)),
                                        as.data.frame(newx)), 
                           formula = Hist(d.time, death != 0) ~ 1, summary = NULL, 
                           se.fit = FALSE, metrics = "brier", contrasts = FALSE, times = times)
#brierFinalResults$models<-as.character(brierFinalResults$models)
# Combine Brier scores between unpenalized and penalized models before plotting

data_brier <- bind_rows(
  brierFinalResults$Brier$score,
  filter(brierFull$Brier$score, model == "PredictionModel")
  )
data_brier$model<-factor(data_brier$model,c("Null model", "PredictionModel","Penalized Cox", "Penalized Logistic"))
levels(data_brier$model)<-c("Kaplan-Meier","Cox","Penalized Cox","Penalized logistic")

reg_brier <- ggplot(data = data_brier, mapping = aes(x = times, y = Brier, col = model)) +
  geom_line() +
  xlab("Follow-up time (Years)") +
  ylab("Brier score") +
  # ggtitle("Brier score over time") +
  labs(color = "Models") +
  paper_gg_theme +  
  scale_color_manual(values = q4)

@@

These values do not inform us of model performance. To do so, we take our hazard functions and use them to predict the absolute risk of everyone in our test set, over time. The probabilities over time for each individual are averaged, resulting in the absolute risk curves in Figure \ref{fig:cs3FinalBrier} (A). There is separation between the curves, but it is minimal. To better demonstrate the differences in performance between the models, we compare Brier scores over time. In Figure \ref{fig:cs3FinalBrier} (B), the Brier score is lower in our penalized logistic regression model compared to both Cox regression models. The differences, however, are minimal. All three outperform the Kaplan-Meier model.

### 2020-08-25 analysis/paper/paper.tex
added reason for smoothish cox and recolored

\caption{\label{tab:bmtcrr-cis}Estimates and confidence intervals for the hazard ratios for each coefficient. Both estimates from case-base sampling and Cox regression are presented.}
\centering
\begin{tabular}[t]{lrlrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Case-Base} & \multicolumn{2}{c}{Cox} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Covariates & HR & 95\% CI & HR & 95\% CI\\
\midrule
Sex & 0.72 & (0.42, 1.26) & 0.68 & (0.39, 1.19)\\
Disease & 0.54 & (0.3, 0.97) & 0.52 & (0.29, 0.93)\\
Phase (CR2 vs. CR1) & 1.21 & (0.48, 3.02) & 1.21 & (0.48, 3.02)\\
Phase (CR3 vs. CR1) & 1.63 & (0.41, 6.44) & 1.67 & (0.43, 6.5)\\
Phase (Relapse vs. CR1) & 4.43 & (2.06, 9.55) & 4.55 & (2.09, 9.9)\\
\addlinespace
Source & 1.51 & (0.49, 4.69) & 1.46 & (0.47, 4.54)\\
Age & 0.99 & (0.97, 1.01) & 0.99 & (0.97, 1.02)\\
\bottomrule
\end{tabular}
\end{table}

@@

Figure \ref{fig:compAbsrisk} shows the absolute risk curves for both
case-base sampling and the Fine-Gray model. As we can see, the two
approaches agree quite well for AML; however, there seems to be a
difference of about 5\% between the two curves for ALL. This difference
does not appear to be significant: the curve from case-base sampling is
contained within a 95\% confidence band around the Fine-Gray absolute
risk curve (figure not shown).

\hypertarget{case-study-3support-data}{%
\section{Case study 3---SUPPORT Data}\label{case-study-3support-data}}

@@

For our penalized model, we opt for the natural log of time; recall that
this corresponds to a Weibull distribution. For fitting the penalized
hazard, we use \code{fitSmoothHazard.fit}, which is a matrix interface
to the function \code{fitSmoothHazard}. In order to use this interface,
we first create a matrix \code{y} containing the time and event
variables, and a matrix \code{x} containing all remaining covariates we
would like to include in the model space. We used lasso penalization by
setting \code{alpha = 1}. Moreover, we use the argument
\code{penalty.factor} to force the time variable into the final model.

@@

Next, we can use the object \code{cbFull} to estimate the absolute risk
curve on the test data. We first create a matrix \code{newx} containing
the relevant variables. Next, using \code{times}, we specify the time
points at which the cumulative incidence will be computed. Finally,
using the parameter \code{s = "lambda.1se"}, we specify at which value
of the tuning parameter we want to retrieve the coefficient estimates.

@@

These values do not inform us of model performance. To do so, we take
our hazard functions and use them to predict the absolute risk of
everyone in our test set, over time. The probabilities over time for
each individual are averaged, resulting in the absolute risk curves in
Figure \ref{fig:cs3FinalBrier} (A). There is separation between the
curves, but it is minimal. To better demonstrate the differences in
performance between the models, we compare Brier scores over time. In
Figure \ref{fig:cs3FinalBrier} (B), the Brier score is lower in our
penalized case-base model compared to both Cox regression models. The
differences, however, are minimal. All three outperform the Kaplan-Meier
estimator.

@@

\caption{\label{fig:cs3FinalBrier} Comparison of Cox regression, penalized Cox regression, penalized case-base estimation, and Kaplan-Meier. (A)  Probabilities of death as it changes over follow-up time. (B) Brier scores as they change over time, where a lower score corresponds to better performace. Penalized case-base estimation produced the lowest results over follow-up time.}\label{fig:riskregressionBrier}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Model1 Model2 Model3 
#>    827    793    792
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> [1] "absRiskCB" "matrix"
\end{CodeOutput}

@@

This report was generated on 2020-08-20 19:21:21 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.2 (2019-12-12)
#> Platform: x86_64-pc-linux-gnu (64-bit)
#> Running under: Pop!_OS 19.10
#> 
#> Matrix products: default
#> BLAS:   /usr/lib/x86_64-linux-gnu/openblas/libblas.so.3
#> LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.3.7.so
#> 
#> locale:
#>  [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
#>  [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
#>  [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
#>  [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
#>  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
#> [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] scales_1.1.1              forcats_0.5.0            
#>  [3] stringr_1.4.0             dplyr_1.0.0              
#>  [5] purrr_0.3.4               readr_1.3.1              
#>  [7] tidyr_1.1.0               ggplot2_3.3.2.9000       
#>  [9] tidyverse_1.3.0           lubridate_1.7.4          
#> [11] visreg_2.7.0              tibble_3.0.1             
#> [13] riskRegression_2020.02.05 prodlim_2019.11.13       
#> [15] flexsurv_1.1.1            survival_3.1-8           
#> [17] pracma_2.2.9              magrittr_1.5             
#> [19] kableExtra_1.1.0          glmnet_4.0-2             
#> [21] Matrix_1.2-18             data.table_1.12.8        
#> [23] cowplot_1.0.0             colorspace_1.4-1         
#> [25] casebase_0.9.0           
#> 
#> loaded via a namespace (and not attached):
#>  [1] TH.data_1.0-10      VGAM_1.1-3          ellipsis_0.3.1     
#>  [4] htmlTable_1.13.3    base64enc_0.1-3     fs_1.3.2           
#>  [7] rstudioapi_0.11     MatrixModels_0.4-1  fansi_0.4.1        
#> [10] mvtnorm_1.0-12      xml2_1.3.0          codetools_0.2-16   
#> [13] knitr_1.29          Formula_1.2-3       jsonlite_1.7.0     
#> [16] broom_0.7.0         cluster_2.1.0       dbplyr_1.4.2       
#> [19] png_0.1-7           compiler_3.6.2      httr_1.4.1         
#> [22] backports_1.1.8     assertthat_0.2.1    cli_2.0.2          
#> [25] acepack_1.4.1       htmltools_0.5.0     quantreg_5.54      
#> [28] tools_3.6.2         gtable_0.3.0        glue_1.4.1         
#> [31] Rcpp_1.0.4.6        cellranger_1.1.0    vctrs_0.3.1        
#> [34] nlme_3.1-143        iterators_1.0.12    xfun_0.15          
#> [37] rvest_0.3.5         lifecycle_0.2.0     polspline_1.1.17   
#> [40] muhaz_1.2.6.1       MASS_7.3-51.5       zoo_1.8-7          
#> [43] hms_0.5.3           sandwich_2.5-1      SparseM_1.78       
#> [46] RColorBrewer_1.1-2  rticles_0.13        yaml_2.2.1         
#> [49] gridExtra_2.3       rms_5.1-4           rpart_4.1-15       
#> [52] latticeExtra_0.6-29 stringi_1.4.6       foreach_1.5.0      
#> [55] checkmate_2.0.0     lava_1.6.6          shape_1.4.4        
#> [58] rlang_0.4.6         pkgconfig_2.0.3     evaluate_0.14      
#> [61] lattice_0.20-38     htmlwidgets_1.5.1   cmprsk_2.2-9       
#> [64] tidyselect_1.1.0    deSolve_1.27.1      R6_2.4.1           
#> [67] generics_0.0.2      Hmisc_4.3-1         multcomp_1.4-12    
#> [70] DBI_1.1.0           withr_2.2.0         pillar_1.4.4       
#> [73] haven_2.3.1         foreign_0.8-74      mgcv_1.8-31        
#> [76] nnet_7.3-12         mstate_0.2.12       modelr_0.1.5       
#> [79] crayon_1.3.4        rmarkdown_2.3       timereg_1.9.4      
#> [82] jpeg_0.1-8.1        grid_3.6.2          readxl_1.3.1       
#> [85] reprex_0.3.0        digest_0.6.25       webshot_0.5.2      
#> [88] numDeriv_2016.8-1.1 stats4_3.6.2        munsell_0.5.0      
#> [91] viridisLite_0.3.0   quadprog_1.5-8
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    fix-cs1 /home/sahir/git_repositories/cbpaper
#> Remote:   fix-cs1 @ origin (https://github.com/sahirbhatnagar/cbpaper.git)
#> Head:     [f618e21] 2020-08-18: time dependent hazard ratio for ERSPC
\end{CodeOutput}
\end{CodeChunk}

>>>>

\caption{\label{tab:bmtcrr-cis}Estimates and confidence intervals for the hazard ratios for each coefficient. Both estimates from case-base sampling and Cox regression are presented.}
\centering
\begin{tabular}[t]{lrlrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Case-Base} & \multicolumn{2}{c}{Cox} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Covariates & HR & 95\% CI & HR & 95\% CI\\
\midrule
Sex & 0.70 & (0.4, 1.23) & 0.68 & (0.39, 1.19)\\
Disease & 0.53 & (0.29, 0.95) & 0.52 & (0.29, 0.93)\\
Phase (CR2 vs. CR1) & 1.12 & (0.45, 2.78) & 1.21 & (0.48, 3.02)\\
Phase (CR3 vs. CR1) & 1.45 & (0.37, 5.66) & 1.67 & (0.43, 6.5)\\
Phase (Relapse vs. CR1) & 4.04 & (1.88, 8.68) & 4.55 & (2.09, 9.9)\\
\addlinespace
Source & 1.48 & (0.48, 4.54) & 1.46 & (0.47, 4.54)\\
Age & 0.99 & (0.97, 1.01) & 0.99 & (0.97, 1.02)\\
\bottomrule
\end{tabular}
\end{table}

@@

Figure \ref{fig:compAbsrisk} shows the absolute risk curves for both
case-base sampling and the Fine-Gray model. As we can see, the two
approaches agree quite well for AML; however, there seems to be a
difference of about 5\% between the two curves for ALL. This difference
does not appear to be significant: the curve from case-base sampling is
contained within a 95\% confidence band around the Fine-Gray absolute
risk curve (figure not shown). \# Case study 3---SUPPORT Data

@@

For our penalized logistic regression model, we opt for the natural log
of time; recall that this corresponds to a Weibull distribution. For
fitting the penalized hazard, we use \code{fitSmoothHazard.fit}, which
is a matrix interface to the function \code{fitSmoothHazard}. In order
to use this interface, we first create a matrix \code{y} containing the
time and event variables, and a matrix \code{x} containing all remaining
covariates we would like to include in the model space. We used lasso
penalization by setting \code{alpha = 1}. Moreover, we use the argument
\code{penalty.factor} to force the time variable into the final model.

@@

Next, we can use the object \code{cbFull} to estimate the absolute risk
curve on the test data. We first create a matrix \code{newx} containing
the relevant variables. Using \code{times}, we specify the time points
at which the cumulative incidence will be computed. Finally, using the
parameter \code{s = "lambda.1se"}, we specify at which value of the
tuning parameter we want to retrieve the coefficient estimates.

@@

@@

\caption{\label{fig:cs3FinalBrier} (A) Four curves comparing the probabilities of death as it changes over follow-up time between Cox regression, penalized Cox regression, penalized logistic regression, and Kaplan-Meier. All follow-up times seen in the new data were used to generate the curves. The Cox-based curves appear to be smooth, but this is due to the 1703 follow-up times seen while fitting the models and the 319 time points used to estimate the curves. (B) Four curves demonstrating the Brier scores as they change over time between Cox regression, penalized Cox regression, penalized logistic regression, and Kaplan-Meier. Lower is better. Penalized logistic regression produced the lowest Brier scores over follow-up time.}\label{fig:riskregressionBrier}
\end{figure}
\end{CodeChunk}

These values do not inform us of model performance. To do so, we take
our hazard functions and use them to predict the absolute risk of
everyone in our test set, over time. The probabilities over time for
each individual are averaged, resulting in the absolute risk curves in
Figure \ref{fig:cs3FinalBrier} (A). There is separation between the
curves, but it is minimal. To better demonstrate the differences in
performance between the models, we compare Brier scores over time. In
Figure \ref{fig:cs3FinalBrier} (B), the Brier score is lower in our
penalized logistic regression model compared to both Cox regression
models. The differences, however, are minimal. All three outperform the
Kaplan-Meier model.

@@

\begin{CodeOutput}
#> Model1 Model2 Model3 
#>    827    790    788
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> [1] "absRiskCB" "matrix"    "array"
\end{CodeOutput}

@@

This report was generated on 2020-08-25 12:59:22 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 4.0.2 (2020-06-22)
#> Platform: x86_64-pc-linux-gnu (64-bit)
#> Running under: Pop!_OS 20.04 LTS
#> 
#> Matrix products: default
#> BLAS:   /usr/lib/x86_64-linux-gnu/blas/libblas.so.3.9.0
#> LAPACK: /usr/lib/x86_64-linux-gnu/lapack/liblapack.so.3.9.0
#> 
#> locale:
#>  [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
#>  [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8    
#>  [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=en_US.UTF-8   
#>  [7] LC_PAPER=en_US.UTF-8       LC_NAME=C                 
#>  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
#> [11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C       
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.7.0               dotwhisker_0.5.0         
#>  [3] timereg_1.9.6             glue_1.4.1               
#>  [5] scales_1.1.1              forcats_0.5.0            
#>  [7] stringr_1.4.0             dplyr_1.0.1              
#>  [9] purrr_0.3.4               readr_1.3.1              
#> [11] tidyr_1.1.1               ggplot2_3.3.2            
#> [13] tidyverse_1.3.0           lubridate_1.7.9          
#> [15] visreg_2.7.0              tibble_3.0.3             
#> [17] riskRegression_2020.06.25 prodlim_2019.11.13       
#> [19] flexsurv_1.1.1            survival_3.2-3           
#> [21] pracma_2.2.9              magrittr_1.5             
#> [23] kableExtra_1.1.0          glmnet_4.0-2             
#> [25] Matrix_1.2-18             data.table_1.13.0        
#> [27] cowplot_1.0.0             colorspace_1.4-1         
#> [29] casebase_0.9.0           
#> 
#> loaded via a namespace (and not attached):
#>  [1] TH.data_1.0-10      VGAM_1.1-3          ellipsis_0.3.1     
#>  [4] ggstance_0.3.4      htmlTable_2.0.1     fs_1.4.2           
#>  [7] base64enc_0.1-3     rstudioapi_0.11     farver_2.0.3       
#> [10] MatrixModels_0.4-1  fansi_0.4.1         mvtnorm_1.1-1      
#> [13] xml2_1.3.2          codetools_0.2-16    knitr_1.29         
#> [16] Formula_1.2-3       jsonlite_1.7.0      cluster_2.1.0      
#> [19] dbplyr_1.4.4        png_0.1-7           compiler_4.0.2     
#> [22] httr_1.4.2          backports_1.1.8     assertthat_0.2.1   
#> [25] cli_2.0.2           acepack_1.4.1       htmltools_0.5.0    
#> [28] quantreg_5.61       tools_4.0.2         gtable_0.3.0       
#> [31] Rcpp_1.0.5          cellranger_1.1.0    vctrs_0.3.2        
#> [34] nlme_3.1-148        conquer_1.0.1       iterators_1.0.12   
#> [37] xfun_0.16           rvest_0.3.6         lifecycle_0.2.0    
#> [40] polspline_1.1.19    muhaz_1.2.6.1       MASS_7.3-51.6      
#> [43] zoo_1.8-8           hms_0.5.3           sandwich_2.5-1     
#> [46] SparseM_1.78        RColorBrewer_1.1-2  rticles_0.14       
#> [49] yaml_2.2.1          gridExtra_2.3       rms_6.0-1          
#> [52] rpart_4.1-15        latticeExtra_0.6-29 stringi_1.4.6      
#> [55] foreach_1.5.0       checkmate_2.0.0     lava_1.6.7         
#> [58] shape_1.4.4         mets_1.2.7.1        rlang_0.4.7        
#> [61] pkgconfig_2.0.3     matrixStats_0.56.0  evaluate_0.14      
#> [64] lattice_0.20-41     labeling_0.3        htmlwidgets_1.5.1  
#> [67] cmprsk_2.2-10       tidyselect_1.1.0    deSolve_1.28       
#> [70] R6_2.4.1            generics_0.0.2      Hmisc_4.4-0        
#> [73] multcomp_1.4-13     DBI_1.1.0           withr_2.2.0        
#> [76] pillar_1.4.6        haven_2.3.1         foreign_0.8-80     
#> [79] mgcv_1.8-31         nnet_7.3-14         mstate_0.2.12      
#> [82] modelr_0.1.8        crayon_1.3.4        rmarkdown_2.3      
#> [85] jpeg_0.1-8.1        grid_4.0.2          readxl_1.3.1       
#> [88] blob_1.2.1          reprex_0.3.0        digest_0.6.25      
#> [91] webshot_0.5.2       numDeriv_2016.8-1.1 stats4_4.0.2       
#> [94] munsell_0.5.0       viridisLite_0.3.0   quadprog_1.5-8
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    master /home/jesseislam/Documents/gits/cbpaper
#> Remote:   master @ origin (git@github.com:Jesse-Islam/cbpaper.git)
#> Head:     [ffe8b03] 2020-08-25: Merge branch 'master' into sbmaster
\end{CodeOutput}
\end{CodeChunk}
